---
title: "Base Model Function"
author: "Elisabeth Nelson"
date: "2022-12-06"
output: pdf_document
---

## Load Packages
```{r}

library(MCMCpack)
library(stats)
library(CARBayes) # <- this is the package for the model
library(rgdal) # <- for working with shapefiles
library(RColorBrewer) 
library(ggplot2)
library(rgeos) # <- also for working with shapefiles
library(maptools) # shapefiles again
library(spdep) # still shapefiles 
library(ggmap) # plotting shapefiles 
library(sf)
library(dplyr)
library(tidyverse)
library(CARBayesST)
library(TSDT)
library(lubridate)

```

## Load Needed Files
```{r}
base_model_final <- readRDS("base_model_final.RDS")
neighb.mat_2 <- readRDS("neighb.mat_2.RDS")

```

## Function for generating dataset
```{r}

#initially, fit through nov 2006, set dec 2006 to NA
#'2006-12-01'

base_model_final_2 <- base_model_final %>%
  mutate(year = as_factor(year)) %>%
  filter(VARNAME_2 != "Kien Hai", 
         VARNAME_2 != "Phu Quoc") %>%
  distinct(year, month, VARNAME_2, NAME_1, ENGTYPE_2, .keep_all = T)


advance_missing_func <- function(N_advance){
      
      base_model_for_model_1 <- base_model_final_2 %>% 
        mutate(year = unfactor(year),
               #monthN = as.numeric(gsub('m','', month)),
               date = as.Date(paste(year, month, '01', sep='-')),
               pop=if_else(is.na(pop), 0.5 ,pop)) %>%
        filter( date <= (as.Date('2006-12-01' ) %m+% months(N_advance))) %>%
        arrange(VARNAME_2, date) %>%
        group_by(VARNAME_2) %>%
        mutate( m_DHF_cases=if_else(is.na(m_DHF_cases),0, m_DHF_cases), ##ALERT...assuming nov, dec missing=0
          index= row_number() , 
               m_DHF_cases_fit = if_else(index==max(index,na.rm=T), NA_real_,m_DHF_cases)) %>%
        dplyr::select(month, year, VARNAME_2, m_DHF_cases,m_DHF_cases_fit, pop) %>%
        arrange(month, year, VARNAME_2)
}

# function for generating dataset without missing month
ds_for_mod_plots_func <- function(N_advance){
      
      base_model_for_model_1 <- base_model_final %>% 
        mutate(year = TSDT::unfactor(year),
               date = as.Date(paste(year, month, '01', sep='-')),
               pop=if_else(is.na(pop), 0.5 ,pop)) %>%
        filter( date <= (as.Date('2006-12-01' ) %m+% months(N_advance) )) %>%
        mutate( m_DHF_cases=if_else(is.na(m_DHF_cases),0, m_DHF_cases)) %>%
        dplyr::select(month, year, VARNAME_2, m_DHF_cases, pop, geometry) %>%
        arrange(month, year, VARNAME_2)
}

ds_mod_aug_2007 <- ds_for_mod_plots_func(8)

```

## Base Model
For this model, n_advance = model # - 1 because need to start at 0 for month 12, 2006
```{r}
###############################################################

#Model Fitting

###############################################################

# default of burnin = 10000, n.sample = 110000, thin = 10

samples_to_quant <- function(X){
  quant1 <- apply(X, 2, quantile, probs=c(0.025, 0.975,seq(0.01, 0.99, by=0.01)))
  return(quant1)
} 



mod.func <- function(N_advance) {
  ds <- advance_missing_func(N_advance) 
 
  x_model = model.matrix(~ as.factor(as.numeric(month)), data = ds)

  offset_model = log(ds$pop) 


  results<-ST.CARar(m_DHF_cases_fit ~ -1 + x_model + offset(offset_model) , 
                  data = ds,

                  family = "poisson",

                  W = neighb.mat_2,

                  burnin = 10000, 

                  n.sample = 110000,

                  thin = 10,

                  prior.mean.beta = rep(0,

                                        times = ncol(x_model)),
                                        
                  prior.var.beta = rep((100^2),

                                       times = ncol(x_model)),

                  prior.tau2 = c(0.01, 0.01),
                  
                  AR = 1,

                  MALA = TRUE,

                  verbose = TRUE)
 #fitted <- results$samples$fitted
 #attributes(fitted) <- NULL
 #posterior_quant <- matrix(fitted)
# posterior_quant <- as.data.frame(fitted)
 posterior_quant_df <- lapply(X = results$samples,FUN = samples_to_quant)
 # save to a file
 saveRDS(posterior_quant_df, paste0("posterior_mod_", N_advance,".rds"))
  return(0)
}

#mod_aug_2007 <- mod.func(8)

 # pull out fitted values
  # fitted <- results$samples$fitted
   # turn into a data frame
  # posterior_quant <- as.data.frame(t(fitted))
   # compute quantiles
   posterior_quant_df <- lapply(X = results$samples$fitted,FUN = samples_to_quant)
   # save to a file
   saveRDS(posterior_quant_df, paste0("posterior_mod_", N_advance,".rds"))
  return(0)
fitted <- mod_aug_2007$samples$fitted
attributes(fitted) <- NULL
posterior_quant <- matrix(fitted)

 posterior_quant_df <- lapply(X = posterior_quant,FUN = samples_to_quant)

mod.func(3)







```


generate summary statistics
take the first percentile --> quantile 1 -100 for each parameter to get the full distribution


## Map against real data
```{r}
library(viridis)
library(scales)

ds_mod_aug_2007$fitted_cases <- mod_aug_2007$fitted.values
#ds_mod_aug_2007$geometry[ds_mod_aug_2007$VARNAME_2 == "Long My Town"] <- missing_long_my

ds_mod_aug_2007 <- ds_mod_aug_2007 %>%
  mutate(real_inc = m_DHF_cases/pop,
         real_inc = real_inc*100000,
         fitted_inc = fitted_cases/pop,
         fitted_inc = fitted_inc*100000,
         log_real_inc = if_else(real_inc == 0, 0, log(real_inc)), 
         log_fitted_inc = log(fitted_inc))

ds_mod_aug_2007 <- st_as_sf(ds_mod_aug_2007)

ds_mod_aug_2007 %>%
  filter(year == 2007 & month == 8) %>%
  ggplot()+
  geom_sf(aes(fill = log_real_inc), lwd = 0)+
  viridis::scale_fill_viridis(option = "viridis", 
                              direction = -1, 
                              name = "Real Inc (log per 100,000)", 
                              oob = scales::squish)+
    theme(axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks = element_blank(),
        panel.background = element_rect(fill = "white", color = NA))

ds_mod_aug_2007 %>%
  filter(year == 2007 & month == 8) %>%
  ggplot()+
  geom_sf(aes(fill = log_fitted_inc), lwd = 0)+
  viridis::scale_fill_viridis(option = "viridis", 
                              direction = -1, 
                              name = "Fitted Inc (log per 100,000)", 
                              oob = scales::squish)+
    theme(axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks = element_blank(),
        panel.background = element_rect(fill = "white", color = NA))
    # make x y plot of observed and fitted values, with reference line with slope of 1, global mean of cases
# mulitply the indcidence by 100,000

ggplot(ds_mod_aug_2007) %>%
  geom_line()

```
could save every 1 percentile of outputs from mcmc on the cluster

## check convergence
```{r}

plot(mod_aug_2007$samples$fitted, type = "l")
head(mod_aug_2007$samples$fitted)

convergence_plot <- mod_aug_2007$samples$fitted

convergence_plot_no_burnin <- convergence_plot

plot(convergence_plot, type = "l", ask=T)

```

```{r}

ds_mod_aug_2007 %>%
  ggplot()+
  geom_point(aes(month, m_DHF_cases), color = "blue")+
  geom_point(aes(month, fitted_cases), color = "red")+
  facet_wrap(~year)

ds_mod_aug_2007 %>%
  ggplot()+
  geom_point(aes(fitted_cases, m_DHF_cases), color = "blue")+
  geom_line(aes(m_DHF_cases, m_DHF_cases), color = "red")+
  #geom_line(aes(fitted_cases, fitted_cases), color = "green")+
  facet_wrap(~year)

ds_mod_aug_2007 %>%
  ggplot()+
  geom_point(aes(fitted_inc, real_inc), color = "blue")+
  geom_line(aes(real_inc, real_inc), color = "red")+
  #geom_line(aes(fitted_inc, fitted_inc), color = "green")+
  facet_wrap(~year)

aug_2007 <- ds_mod_aug_2007 %>%
  filter(year == 2007 & month == 8) %>%
  mutate(fitted_3rd_quant = as.numeric(posterior_quant$Y[3, ]), 
         fitted_97th_quant = as.numeric(posterior_quant$Y[97, ])) %>%
  mutate(fitted_3_inc = (fitted_3rd_quant/pop)*100000, 
         fitted_97_inc = (fitted_97th_quant/pop)*100000, 
         log_fitted_3_inc = log(fitted_3_inc), 
         log_fitted_97_inc = log(fitted_97_inc))

  
ggplot(data = aug_2007)+
  geom_point(aes(fitted_cases, m_DHF_cases), color = "blue")+
  geom_errorbar(aes(fitted_cases, ymin = fitted_3rd_quant, 
                    ymax = fitted_97th_quant), color = "green")+
  geom_line(aes(m_DHF_cases, m_DHF_cases), color = "red")

ggplot(data = aug_2007)+
  geom_point(aes(log_fitted_inc, log_real_inc), color = "blue")+
    geom_errorbar(aes(log_fitted_inc, ymin = log_fitted_3_inc, 
                    ymax = log_fitted_97_inc), color = "green")+
  geom_line(aes(log_real_inc, log_real_inc), color = "red")

model_fit <- ds_mod_aug_2007 %>%
  mutate(fitted_3rd_quant = as.numeric(posterior_quant$Y[3, ]), 
         fitted_97th_quant = as.numeric(posterior_quant$Y[97, ])) %>%
  mutate(fitted_3_inc = (fitted_3rd_quant/pop)*100000, 
         fitted_97_inc = (fitted_97th_quant/pop)*100000, 
         log_fitted_3_inc = log(fitted_3_inc), 
         log_fitted_97_inc = log(fitted_97_inc))

```

use 8th month  in 2007
sqrt x and y variables 
compare to 95% CI of y

do for all timepoints once run through the cluster

mean absolute error