---
title: "Climate_Model_Function"
author: "Elisabeth Nelson"
date: "2023-03-01"
output: pdf_document
---

## Load Packages
```{r}

library(MCMCpack)
library(stats)
library(CARBayes) # <- this is the package for the model
library(rgdal) # <- for working with shapefiles
library(RColorBrewer) 
library(ggplot2)
library(rgeos) # <- also for working with shapefiles
library(maptools) # shapefiles again
library(spdep) # still shapefiles 
library(ggmap) # plotting shapefiles 
library(sf)
library(dplyr)
library(tidyverse)
library(CARBayesST)
library(TSDT)
library(lubridate)

```

## Load Climate Data
name,units,long name
u10,m s**-1,10 metre U wind component
v10,m s**-1,10 metre V wind component
d2m,K,2 metre dewpoint temperature
t2m,K,2 metre temperature
lsm,(0 - 1),Land-sea mask
tp,m,Total precipitation
sp,Pa,Surface pressure
d2m,K,dew point temperature
rh,%,relative temperature
```{r}
#district_list <- c("019")
#district_list <- c("073")
#district_list <- c("051", "072", "000", "054", "099", "086", "022", "062", "094")
#district_list <- c("002", "028", "033", "035", "064", "065", "068", "080", "082", "093")
#district_list <- c("078", "092", "095", "031", "009", "030", "056", "057", "083", "090")
#district_list <- c("006", "034", "096", "098", "059", "085", "003", "004", "063", "087")
#district_list <- c("007", "021", "005", "045", "069", "091", "018", "037", "044", "048")
#district_list <- c("011", "029", "041", "077", "032", "046", "049", "050", "089", "053")
#district_list <- c("043", "070", "017", "042", "040", "008", "012", "036", "075", "066")
#district_list <- c("088", "084", "001", "076", "071", "097", "026", "079", "015", "067")
#district_list <- c("081", "024", "039", "055", "074", "025", "058", "016", "027", "061")
#district_list <- c("010", "014", "020", "038", "052", "060", "013", "023", "132", "133")
#district_list <- c("125", "047", "126", "127", "128", "129", "130", "131", "113", "114")
#district_list <- c("123", "124", "119", "120", "121", "122", "115", "116", "117", "118")
#district_list <- c("101", "109", "102", "105", "106", "107", "100", "103", "104", "108")
district_list <- c("110", "111", "112")

for(i in district_list){
  N_district <- i
  district.climate.func <- function(N_district){
    
    # import district climate files 
      district <- do.call(rbind, 
                         lapply(list.files(paste0('/Users/zoe/Documents/Yale EMD/Rotation 1/Dengue - Vietnam/dengue_exploration/Data/climate/', N_district), 
                         recursive = TRUE, full.names = TRUE), read_csv))

    # calculate monthly averages (and total precipitation)
      district <- district %>%
        mutate(district = N_district, 
               month = month(time), 
               year = year(time)) %>%
        group_by(month, year) %>%
        summarize(u10 = mean(u10), 
                  v10 = mean(v10), 
                  d2m = mean(d2m), 
                  t2m = mean(t2m), 
                  lsm = mean(lsm), 
                  tp = sum(tp), 
                  sp = mean(sp), 
                  d2m = mean(d2m), 
                  rh = mean(rh))
      
    # save as an RDS file
      saveRDS(district, paste0("Data/climate/district_climate/district", N_district, ".rds"))
  }
 district.climate.func(i) 
}
```

## Combine district files 
```{r}
## fix data column
district_list <- c("019", "073", "051", "072", "000", "054", "099", "086", "022", "062", "094",
                   "002", "028", "033", "035", "064", "065", "068", "080", "082", "093",
                   "078", "092", "095", "031", "009", "030", "056", "057", "083", "090",
                   "006", "034", "096", "098", "059", "085", "003", "004", "063", "087",
                   "007", "021", "005", "045", "069", "091", "018", "037", "044", "048",
                   "011", "029", "041", "077", "032", "046", "049", "050", "089", "053",
                   "043", "070", "017", "042", "040", "008", "012", "036", "075", "066",
                   "088", "084", "001", "076", "071", "097", "026", "079", "015", "067",
                   "081", "024", "039", "055", "074", "025", "058", "016", "027", "061",
                   "010", "014", "020", "038", "052", "060", "013", "023", "132", "133",
                   "125", "047", "126", "127", "128", "129", "130", "131", "113", "114",
                   "123", "124", "119", "120", "121", "122", "115", "116", "117", "118",
                   "101", "109", "102", "105", "106", "107", "100", "103", "104", "108",
                   "110", "111", "112")

for(i in district_list){
  N_district <- i 
  district.column.func <- function(N_district) {
    # read in district climate file
    district <- readRDS(paste0('/Users/zoe/Documents/Yale EMD/Rotation 1/Dengue - Vietnam/dengue_exploration/Data/climate/district_climate/district', N_district, '.rds'))
    # add district column
    district <- as.data.frame(district) %>%
      mutate(district_num = N_district)
    # save as an RDS file
    saveRDS(district, paste0("Data/climate/district_climate_1/district", N_district, ".rds"))
  }
  district.column.func(i)
}

## combine all files
monthly_climate <- do.call(rbind, 
                         lapply(list.files('/Users/zoe/Documents/Yale EMD/Rotation 1/Dengue - Vietnam/dengue_exploration/Data/climate/district_climate_1', 
                         recursive = TRUE, full.names = TRUE), readRDS))

monthly_climate <- monthly_climate %>%
  mutate(district_num = as.numeric(district_num))

## import district code df
district_codes <- read_csv("Data/climate/readme.districts.txt", 
    col_types = cols(...1 = col_number()))
district_codes <- as.data.frame(district_codes)

## combine district codes with
monthly_climate_final <- left_join(monthly_climate, district_codes, by = c("district_num" = "...1"))

```

## Clean data to fit to base model 
```{r}

monthly_climate_final_2 <- monthly_climate_final %>%
  filter(NAME_1 == "An Giang" & VARNAME_2 == "Chau Thanh") %>%
  mutate(VARNAME_2 = str_replace(VARNAME_2, "Chau Thanh", "Chau Thanh An Giang"))

monthly_climate_final_3 <- monthly_climate_final %>%
  filter(NAME_1 == "Bến Tre" & VARNAME_2 == "Chau Thanh") %>%
  mutate(VARNAME_2 = str_replace(VARNAME_2, "Chau Thanh", "Chau Thanh Ben Tre"))

monthly_climate_final_4 <- monthly_climate_final %>%
  filter(NAME_1 == "Đồng Tháp" & VARNAME_2 == "Chau Thanh") %>%
  mutate(VARNAME_2 = str_replace(VARNAME_2, "Chau Thanh", "Chau Thanh Dong Thanp"))

monthly_climate_final_5 <- monthly_climate_final %>%
  filter(NAME_1 == "Hậu Giang" & VARNAME_2 == "Chau Thanh") %>%
  mutate(VARNAME_2 = str_replace(VARNAME_2, "Chau Thanh", "Chau Thanh Hau Giang"))

monthly_climate_final_6 <- monthly_climate_final %>%
  filter(NAME_1 == "Kiên Giang" & VARNAME_2 == "Chau Thanh") %>%
  mutate(VARNAME_2 = str_replace(VARNAME_2, "Chau Thanh", "Chau Thanh Kien Giang"))

monthly_climate_final_7 <- monthly_climate_final %>%
  filter(NAME_1 == "Long An" & VARNAME_2 == "Chau Thanh") %>%
  mutate(VARNAME_2 = str_replace(VARNAME_2, "Chau Thanh", "Chau Thanh Long An"))
 
monthly_climate_final_8 <- monthly_climate_final %>%
  filter(NAME_1 == "Sóc Trăng" & VARNAME_2 == "Chau Thanh") %>%
  mutate(VARNAME_2 = str_replace(VARNAME_2, "Chau Thanh", "Chau Thanh Soc Trang")) 
         
monthly_climate_final_9 <- monthly_climate_final %>%
  filter(NAME_1 == "Tiền Giang" & VARNAME_2 == "Chau Thanh") %>%
  mutate(VARNAME_2 = str_replace(VARNAME_2, "Chau Thanh", "Chau Thanh Tien Giang"))

monthly_climate_final_10 <- monthly_climate_final %>%
  filter(NAME_1 == "Trà Vinh" & VARNAME_2 == "Chau Thanh") %>%
  mutate(VARNAME_2 = str_replace(VARNAME_2, "Chau Thanh", "Chau Thanh Tra Vinh"))

monthly_climate_final_11 <- monthly_climate_final %>%
  filter(NAME_1 == "An Giang" & VARNAME_2 == "Phu Tan") %>%
  mutate(VARNAME_2 = str_replace(VARNAME_2, "Phu Tan", "Phu Tan An Giang")) 
      
monthly_climate_final_12 <- monthly_climate_final %>%
  filter(NAME_1 == "Cà Mau" & VARNAME_2 == "Phu Tan") %>%
  mutate(VARNAME_2 = str_replace(VARNAME_2, "Phu Tan", "Phu Tan Ca Mau"))

monthly_climate_final_13 <- monthly_climate_final %>%
  filter(NAME_1 == "Đồng Tháp" & VARNAME_2 == "Cao Lanh" & ENGTYPE_2 == "District") %>%
  mutate(VARNAME_2 = str_replace(VARNAME_2, "Cao Lanh", "Cao Lanh District"))

monthly_climate_final_14 <- monthly_climate_final %>%
  filter(NAME_1 == "Đồng Tháp" & VARNAME_2 == "Cao Lanh" & ENGTYPE_2 == "City") %>%
  mutate(VARNAME_2 = str_replace(VARNAME_2, "Cao Lanh", "Cao Lanh City"))

monthly_climate_final_15 <- monthly_climate_final %>%
  filter(NAME_1 == "Đồng Tháp" & VARNAME_2 == "Hong Ngu" & ENGTYPE_2 == "District") %>%
  mutate(VARNAME_2 = str_replace(VARNAME_2, "Hong Ngu", "Hong Ngu District"))

monthly_climate_final_16 <- monthly_climate_final %>%
  filter(NAME_1 == "Đồng Tháp" & VARNAME_2 == "Hong Ngu" & ENGTYPE_2 == "City") %>%
  mutate(VARNAME_2 = str_replace(VARNAME_2, "Hong Ngu", "Hong Ngu City"))

monthly_climate_final_17 <- monthly_climate_final %>%
  filter(NAME_1 == "Hậu Giang" & VARNAME_2 == "Long My" & ENGTYPE_2 == "District") %>%
  mutate(VARNAME_2 = str_replace(VARNAME_2, "Long My", "Long My District"))

monthly_climate_final_18 <- monthly_climate_final %>%
  filter(NAME_1 == "Hậu Giang" & VARNAME_2 == "Long My" & ENGTYPE_2 == "Town") %>%
  mutate(VARNAME_2 = str_replace(VARNAME_2, "Long My", "Long My Town"))

monthly_climate_final_19 <- monthly_climate_final %>%
  filter(NAME_1 == "Tiền Giang" & VARNAME_2 == "Cai Lay" & ENGTYPE_2 == "District") %>%
  mutate(VARNAME_2 = str_replace(VARNAME_2, "Cai Lay", "Cai Lay District"))

monthly_climate_final_20 <- monthly_climate_final %>%
  filter(NAME_1 == "Tiền Giang" & VARNAME_2 == "Cai Lay" & ENGTYPE_2 == "Town") %>%
  mutate(VARNAME_2 = str_replace(VARNAME_2, "Cai Lay", "Cai Lay Town"))

monthly_climate_final_21 <- monthly_climate_final %>%
  filter(NAME_1 == "Trà Vinh" & VARNAME_2 == "Duyen Hai" & ENGTYPE_2 == "District") %>%
  mutate(VARNAME_2 = str_replace(VARNAME_2, "Duyen Hai", "Duyen Hai District"))

monthly_climate_final_22 <- monthly_climate_final %>%
  filter(NAME_1 == "Trà Vinh" & VARNAME_2 == "Duyen Hai" & ENGTYPE_2 == "Town") %>%
  mutate(VARNAME_2 = str_replace(VARNAME_2, "Duyen Hai", "Duyen Hai Town"))

monthly_climate_final_filter <- rbind(monthly_climate_final, monthly_climate_final_2, monthly_climate_final_3, monthly_climate_final_4, monthly_climate_final_5, monthly_climate_final_6, monthly_climate_final_7, monthly_climate_final_8, monthly_climate_final_9, monthly_climate_final_10, monthly_climate_final_11, monthly_climate_final_12, monthly_climate_final_13, monthly_climate_final_14, monthly_climate_final_15, monthly_climate_final_16, monthly_climate_final_17, monthly_climate_final_18, monthly_climate_final_19, monthly_climate_final_20, monthly_climate_final_21, monthly_climate_final_22)

names_to_delete <- c("Chau Thanh", "Phu Tan", "Cao Lanh", "Hong Ngu", "Long My", "Cai Lay", "Duyen Hai")

monthly_climate_final_filter <- monthly_climate_final_filter %>%
  filter(!(VARNAME_2 %in% names_to_delete))

## save as RDS
saveRDS(monthly_climate_final_filter, "Data/climate/monthly_climate_data.rds")
```

## Join with base model
```{r}
base_model_final <- readRDS("base_model_final.RDS")
base_model_final <- base_model_final %>%
  mutate(year = as.character(year),
        year = as.numeric(year)) %>%
  distinct(year, month, VARNAME_2, NAME_1, ENGTYPE_2, .keep_all = T)


climate_model <- left_join(base_model_final, monthly_climate_final_filter, by = c("VARNAME_2" = "VARNAME_2", "NAME_1" = "NAME_1", "NAME_2" = "NAME_2", 
                                                                                  "TYPE_2" = "TYPE_2", "ENGTYPE_2" = "ENGTYPE_2", "month" = "month", 
                                                                                  "year" = "year"))
climate_model <- climate_model %>%
  select(!(c(district_num, GID_0, NAME_0, GID_1, NL_NAME_1, GID_2, NL_NAME_2, CC_2, HASC_2)))

temp_model <- climate_model %>%
  group_by(VARNAME_2) %>%
  select(!(c(u10, v10, d2m, t2m, lsm, tp, sp))) %>%
  mutate(lag_1_rh = lag(rh, 1), 
         lag_2_rh = lag(rh, 2), 
         lag_3_rh = lag(rh, 3), 
         lag_4_rh = lag(rh, 4), 
         lag_5_rh = lag(rh, 5), 
         lag_6_rh = lag(rh, 6)) %>%
  mutate(rh = if_else(is.na(rh), 0, rh), 
         lag_1_rh = if_else(is.na(lag_1_rh), 0, lag_1_rh), 
         lag_2_rh = if_else(is.na(lag_2_rh), 0, lag_2_rh), 
         lag_3_rh = if_else(is.na(lag_3_rh), 0, lag_3_rh), 
         lag_4_rh = if_else(is.na(lag_4_rh), 0, lag_4_rh), 
         lag_5_rh = if_else(is.na(lag_5_rh), 0, lag_5_rh), 
         lag_6_rh = if_else(is.na(lag_6_rh), 0, lag_6_rh))
  

ppt_model <- climate_model %>%
  group_by(VARNAME_2) %>%
  select(!(c(u10, v10, d2m, t2m, lsm, sp, rh))) %>%
  mutate(lag_1_tp = lag(tp, 1), 
         lag_2_tp = lag(tp, 2), 
         lag_3_tp = lag(tp, 3), 
         lag_4_tp = lag(tp, 4), 
         lag_5_tp = lag(tp, 5), 
         lag_6_tp = lag(tp, 6)) %>%
  mutate(tp = if_else(is.na(tp), 0, tp), 
         lag_1_tp = if_else(is.na(lag_1_tp), 0, lag_1_tp), 
         lag_2_tp = if_else(is.na(lag_2_tp), 0, lag_2_tp), 
         lag_3_tp = if_else(is.na(lag_3_tp), 0, lag_3_tp), 
         lag_4_tp = if_else(is.na(lag_4_tp), 0, lag_4_tp), 
         lag_5_tp = if_else(is.na(lag_5_tp), 0, lag_5_tp), 
         lag_6_tp = if_else(is.na(lag_6_tp), 0, lag_6_tp))

## save as RDS
saveRDS(climate_model, "Data/climate/climate_model.rds")
saveRDS(temp_model, "Data/climate/temp_model.rds")
saveRDS(ppt_model, "Data/climate/ppt_model.rds")
```

## Load Needed Files
```{r}
base_model_final <- readRDS("base_model_final.RDS")
neighb.mat_2 <- readRDS("neighb.mat_2.RDS")
temp_model <- readRDS("Data/climate/temp_model.rds")
ppt_model <- readRDS("Data/climate/ppt_model.rds")

```

## Function for generating dataset for temp model
```{r}

#initially, fit through nov 2006, set dec 2006 to NA
#'2006-12-01'

temp_model_2 <- temp_model %>%
  mutate(year = as_factor(year)) %>%
  filter(VARNAME_2 != "Kien Hai", 
         VARNAME_2 != "Phu Quoc") %>%
  distinct(year, month, VARNAME_2, NAME_1, ENGTYPE_2, .keep_all = T)


advance_missing_func_temp <- function(N_advance){
      
      temp_model_for_model <- temp_model_2 %>% 
        mutate(year = unfactor(year),
               #monthN = as.numeric(gsub('m','', month)),
               date = as.Date(paste(year, month, '01', sep='-')),
               pop=if_else(is.na(pop), 0.5 ,pop)) %>%
        filter( date <= (as.Date('2006-12-01' ) %m+% months(N_advance))) %>%
        arrange(VARNAME_2, date) %>%
        group_by(VARNAME_2) %>%
        mutate( m_DHF_cases=if_else(is.na(m_DHF_cases),0, m_DHF_cases), ##ALERT...assuming nov, dec missing=0
          index= row_number() , 
               m_DHF_cases_fit = if_else(index==max(index,na.rm=T), NA_real_,m_DHF_cases)) %>%
        dplyr::select(month, year, VARNAME_2, m_DHF_cases,m_DHF_cases_fit, pop, 
                      rh, lag_1_rh, lag_2_rh, lag_3_rh, lag_4_rh, lag_5_rh, lag_6_rh) %>%
        arrange(month, year, VARNAME_2)
}

```

## Function for generating dataset for ppt model
```{r}

#initially, fit through nov 2006, set dec 2006 to NA
#'2006-12-01'

ppt_model_2 <- ppt_model %>%
  mutate(year = as_factor(year)) %>%
  filter(VARNAME_2 != "Kien Hai", 
         VARNAME_2 != "Phu Quoc") %>%
  distinct(year, month, VARNAME_2, NAME_1, ENGTYPE_2, .keep_all = T)


advance_missing_func_ppt <- function(N_advance){
      
      ppt_model_for_model <- ppt_model_2 %>% 
        mutate(year = unfactor(year),
               #monthN = as.numeric(gsub('m','', month)),
               date = as.Date(paste(year, month, '01', sep='-')),
               pop=if_else(is.na(pop), 0.5 ,pop)) %>%
        filter( date <= (as.Date('2006-12-01' ) %m+% months(N_advance))) %>%
        arrange(VARNAME_2, date) %>%
        group_by(VARNAME_2) %>%
        mutate( m_DHF_cases=if_else(is.na(m_DHF_cases),0, m_DHF_cases), ##ALERT...assuming nov, dec missing=0
          index= row_number() , 
               m_DHF_cases_fit = if_else(index==max(index,na.rm=T), NA_real_,m_DHF_cases)) %>%
        dplyr::select(month, year, VARNAME_2, m_DHF_cases,m_DHF_cases_fit, pop, 
                      tp, lag_1_tp, lag_2_tp, lag_3_tp, lag_4_tp, lag_5_tp, lag_6_tp) %>%
        arrange(month, year, VARNAME_2)
}

```

## Fits Model
For this model, n_advance = model # - 1 because need to start at 0 for month 12, 2006
```{r}

# default of burnin = 10000, n.sample = 110000, thin = 10

mod.func <- function(N_advance) {
  ds <- advance_missing_func_temp(N_advance) 
 
  x_model_1 <- model.matrix(~ as.factor(as.numeric(month)), data = ds)
  x_model <- cbind(x_model_1, ds$rh, ds$lag_1_rh, ds$lag_2_rh, ds$lag_3_rh, 
                   ds$lag_4_rh, ds$lag_5_rh, ds$lag_6_rh)

  offset_model <- log(ds$pop) 


  results<-ST.CARar(m_DHF_cases_fit ~ -1 + x_model + offset(offset_model) , 
                  data = ds,

                  family = "poisson",

                  W = neighb.mat_2,

                  burnin = 10, 

                  n.sample = 110,

                  thin = 10,

                  prior.mean.beta = rep(0,

                                        times = ncol(x_model)),
                                        
                  prior.var.beta = rep((100^2),

                                       times = ncol(x_model)),

                  prior.tau2 = c(0.01, 0.01),
                  
                  AR = 1,

                  MALA = TRUE,

                  verbose = TRUE)

  # save fitted values
  post_fit_temp <- results$fitted.values
  saveRDS(post_fit_temp, paste0("post_fit_temp_", N_advance,".rds"))
  return(0)
}

mod.func(2)

```


## Quantiles model
```{r}

samples_to_quant <- function(X){
  quant1 <- apply(X, 2, quantile, probs=c(0.025, 0.975,seq(0.01, 0.99, by=0.01)))
  return(quant1)
} 

mod.func <- function(N_advance) {
  ds <- advance_missing_func(N_advance) 
  
  x_model = model.matrix(~ as.factor(as.numeric(month)), data = ds)
  
  offset_model = log(ds$pop) 
  
  
  results<-ST.CARar(m_DHF_cases_fit ~ -1 + x_model + offset(offset_model) , 
                    data = ds,
                    
                    family = "poisson",
                    
                    W = neighb.mat_2,
                    
                    burnin = 1000, 
                    
                    n.sample = 11000,
                    
                    thin = 10,
                    
                    prior.mean.beta = rep(0.00,
                                          
                                          times = ncol(x_model)),
                    
                    prior.var.beta = rep((100.00^2),
                                         
                                         times = ncol(x_model)),
                    
                    prior.tau2 = c(0.01, 0.01),
                    
                    AR = 1,
                    
                    MALA = FALSE,
                    
                    verbose = TRUE)

  
  # save 1% of fits + 95% CI
  post_quantiles <- lapply(X = results$samples,FUN = samples_to_quant)
  
  # save only fitted values
  post_fitted_quant <- post_quantiles$fitted
  
  # save to a file
  saveRDS(post_fitted_quant, paste0("posterior_quant_", N_advance,".rds"))
  
  return(0)
}

```
