---
title: "Dengue Data Exploration"
author: "Elisabeth Nelson"
date: "2022-10-02"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# set the libraries
library(dplyr)
library(readr)
library(readxl)
library(ggplot2)
library(maptools)
library(spdep)
library(rgeos)
library(sp)
library(MASS)
library(surveillance)
library(tidyr)
library(splines)
library(tidyverse)
library(tsibble)
library(sf)
library(scales)
library(RColorBrewer)
library(viridis)

```


## Import the Data and Combine Incidence and Population Data
```{r}
# import the data

### we only have 132 districts' historical dengue incidence, and only 127 districts' population
    # try to convert all to csv and then look at the last row --> is it super long??????
    # check to make sure the csv looks the same in r (unmbers/decimals)
MDR_Dengue_Cases <- read_csv("Data/CONFIDENTIAL/MDR Dengue Cases.csv")
MDR_district_cases_pop <- read_csv("Data/CONFIDENTIAL/updated - population.csv")

# rename columns and change variable structures
MDR_district_cases_pop <- MDR_district_cases_pop %>%
  rename('no_symbols_name' = 'District') %>%
  rename('old_district_code' = 'District code - Old MDR map') %>%
  rename('updated_district_code' = 'District code - Updated MDR map') %>%
  rename('population' = 'pop') %>%
  mutate(Province = str_replace(Province, "Can Tho city", "Can Tho")) %>%
  mutate(no_symbols_name = tolower(no_symbols_name),
         symbols_name = tolower(symbols_name)) %>%
  dplyr::select(-'...7', -'...8', -'...9', -'...10', -'...11', -'...12')

MDR_Dengue_Cases <- MDR_Dengue_Cases %>%
  mutate(district = tolower(district)) %>%
  mutate(district = str_replace(district, "h. vĩnh lợi", "vĩnh lợi")) %>%
  mutate(district = str_replace(district, "mỏ cày/mỏ cày nam", "mỏ cày nam")) %>%
  mutate(district = str_replace(district, "tx. cao lãnh", "tp. cao lãnh")) %>%
  mutate(district = str_replace(district, "tiểu cần", "tiêu cần")) 

  

updated_MDR_Dengue_Cases <- left_join(MDR_Dengue_Cases, MDR_district_cases_pop, by = c("province" = "Province", "district" = "symbols_name"), na.rm = na.rm)

# fixed mo cay district --> combine populations of mo cay nam and mo cay bac based on updated district codes
# fixed cần thơ not --> combine Binh Thuy, Co Do, Cai Rang, Ninh Kieu, Phong Dien, Thoi Lai based on updated district codes 

  


```

## Changing the District Names
Key for types of districts:
Huyen = District

Thanh pho = City
Chau Doc, Long Xuyen, Bac Lieu, Ben Tre, Ca Mau, Thanh Pho Cao Lanh, Thi xa Hong Ngu, Sa Dec, Nga Bay, Vi Thanh, Ha Tien, Phu Quoc, Rach Gia, Tan An, Soc Trang, My Tho, Tra Vinh, Vinh Long

Thi xa = Town
Tan Chau, Gia Rai, Long My town, Kien Tuong, Nga Nam, Vinh Chau, Thi xa Cai Lay, Go Cong, Duyen Hai town, Binh Minh

Quan = Urban District 
Binh Thuy, Cai Rang, Ninh Kieu, O Mon, Thot Not

```{r}
# Create a column with district names matching VARNAME_2
  
updated_MDR_Dengue_Cases <- updated_MDR_Dengue_Cases %>%
  mutate(no_symbols_name = str_replace(no_symbols_name, "thanh pho cao lanh" , "cao lanh")) %>%
  mutate(no_symbols_name = str_replace(no_symbols_name, "thi xa hong ngu", "hong ngu")) %>%
  mutate(no_symbols_name = str_replace(no_symbols_name, "thi xa cai lay", "cai lay")) %>%
  mutate(no_symbols_name = str_replace(no_symbols_name, "huyen cao lanh", "cao lanh")) %>%
  mutate(no_symbols_name = str_replace(no_symbols_name, "huyen hong ngu", "hong ngu")) %>%
  mutate(no_symbols_name = str_replace(no_symbols_name, "huyen cai lay", "cai lay")) %>%
  mutate(no_symbols_name = str_replace(no_symbols_name, "duyen hai town", "duyen hai")) %>%
  mutate(no_symbols_name = str_replace(no_symbols_name, "long my town", "long my"))

# Types of Districts
City <- c("chau doc", "long xuyen", "bac lieu", "ben tre", "ca mau", "thanh pho cao lanh", "thi xa hong ngu", "sa dec", "nga bay", "vi thanh", "ha tien", "phu quoc", "rach gia", "tan an", "soc trang", "my tho", "tra vinh", "vinh long")
Town <- c("tan chau", "gia rai", "long my town", "kien tuong", "nga nam", "vinh chau", "thi xa cai lay", "go cong", "duyen hai town", "binh minh")
Urban_District <- c("binh thuy", "cai rang", "ninh kieu", "o mon", "thot not")
District <- updated_MDR_Dengue_Cases$no_symbols_name 
District <- District[!District %in% City]
District <- District[!District %in% Town]
District <- District[!District %in% Urban_District]


# Create a column with district types matching ENGTYPE_2

updated_MDR_Dengue_Cases <- updated_MDR_Dengue_Cases %>%
  mutate(district_type = case_when(
    no_symbols_name %in% City ~ "city", 
    no_symbols_name %in% Town ~ "town", 
    no_symbols_name %in% Urban_District ~ "urban district", 
    no_symbols_name %in% District ~ "district"
  )) 
tail(updated_MDR_Dengue_Cases)
```

## Upload Population Estimates
```{r}

pop_2001 <- read_csv("Data/CONFIDENTIAL/updated_populations/2001.csv", 
    col_types = cols(`population(thousand)` = col_number()))
pop_2002 <- read_csv("Data/CONFIDENTIAL/updated_populations/2002.csv", 
    col_types = cols(`population(thousand)` = col_number()))
pop_2003 <- read_csv("Data/CONFIDENTIAL/updated_populations/2003.csv", 
    col_types = cols(`population(thousand)` = col_number()))
pop_2004 <- read_csv("Data/CONFIDENTIAL/updated_populations/2004.csv", 
    col_types = cols(`population(thousand)` = col_number()))
pop_2005 <- read_csv("Data/CONFIDENTIAL/updated_populations/2005.csv", 
    col_types = cols(`population(thousand)` = col_number()))
pop_2006 <- read_csv("Data/CONFIDENTIAL/updated_populations/2006.csv", 
    col_types = cols(`population(thousand)` = col_number()))
pop_2007 <- read_csv("Data/CONFIDENTIAL/updated_populations/2007.csv", 
    col_types = cols(`population(thousand)` = col_number()))
pop_2008 <- read_csv("Data/CONFIDENTIAL/updated_populations/2008.csv", 
    col_types = cols(`population(thousand)` = col_number()))
pop_2009 <- read_csv("Data/CONFIDENTIAL/updated_populations/2009.csv", 
    col_types = cols(`population(thousand)` = col_number()))
pop_2010 <- read_csv("Data/CONFIDENTIAL/updated_populations/2010.csv", 
    col_types = cols(`population(thousand)` = col_number()))
pop_2011 <- read_csv("Data/CONFIDENTIAL/updated_populations/2011.csv", 
    col_types = cols(`population(thousand)` = col_number()))
pop_2012 <- read_csv("Data/CONFIDENTIAL/updated_populations/2012.csv", 
    col_types = cols(`population(thousand)` = col_number()))
pop_2013 <- read_csv("Data/CONFIDENTIAL/updated_populations/2013.csv", 
    col_types = cols(`population(thousand)` = col_number()))
pop_2014 <- read_csv("Data/CONFIDENTIAL/updated_populations/2014.csv", 
    col_types = cols(`population(thousand)` = col_number()))
pop_2015 <- read_csv("Data/CONFIDENTIAL/updated_populations/2015.csv", 
    col_types = cols(`population(thousand)` = col_number()))
pop_2016 <- read_csv("Data/CONFIDENTIAL/updated_populations/2016.csv", 
    col_types = cols(`population(thousand)` = col_number()))
pop_2017 <- read_csv("Data/CONFIDENTIAL/updated_populations/2017.csv", 
    col_types = cols(`population(thousand)` = col_number()))
pop_2018 <- read_csv("Data/CONFIDENTIAL/updated_populations/2018.csv", 
    col_types = cols(`population(thousand)` = col_number()))

# add a column for year
pop_2001 <- pop_2001 %>%
  mutate(year = 2001)
pop_2002 <- pop_2002 %>%
  mutate(year = 2002)
pop_2003 <- pop_2003 %>%
  mutate(year = 2003)
pop_2004 <- pop_2004 %>%
  mutate(year = 2004)
pop_2005 <- pop_2005 %>%
  mutate(year = 2005)
pop_2006 <- pop_2006 %>%
  mutate(year = 2006)
pop_2007 <- pop_2007 %>%
  mutate(year = 2007)
pop_2008 <- pop_2008 %>%
  mutate(year = 2008)
pop_2009 <- pop_2009 %>%
  mutate(year = 2009)
pop_2010 <- pop_2010 %>%
  mutate(year = 2010)
pop_2011 <- pop_2011 %>%
  mutate(year = 2011)
pop_2012 <- pop_2012 %>%
  mutate(year = 2012)
pop_2013 <- pop_2013 %>%
  mutate(year = 2013)
pop_2014 <- pop_2014 %>%
  mutate(year = 2014)
pop_2015 <- pop_2015 %>%
  mutate(year = 2015)
pop_2016 <- pop_2016 %>%
  mutate(year = 2016)
pop_2017 <- pop_2017 %>%
  mutate(year = 2017)
pop_2018 <- pop_2018 %>%
  mutate(year = 2018)

updated_populations <- rbind(pop_2001, pop_2002, pop_2003, pop_2004, pop_2005, pop_2006, pop_2007, pop_2008, pop_2009, pop_2010, pop_2011, pop_2012, pop_2013, pop_2014, pop_2015, pop_2016, pop_2017, pop_2018)

updated_populations_2 <- updated_populations %>%
  dplyr::select(c(VARNAME_2, `population(thousand)`, year)) %>%
  mutate(VARNAME_2 = tolower(VARNAME_2), 
         population = `population(thousand)`*1000)

# download to combine
library(writexl)
write_xlsx(updated_populations_2, "/Users/zoe/Documents/Yale EMD/Rotation 1/Dengue - Vietnam/dengue_exploration\\updated_pop_2.xlsx")
write_xlsx(MDR_2, "/Users/zoe/Documents/Yale EMD/Rotation 1/Dengue - Vietnam/dengue_exploration\\map.xlsx")

# re-upload completed population data
complete_pops <- read_csv("Data/CONFIDENTIAL/updated_pop_2.csv", 
    col_types = cols(`population(thousand)` = col_number(), 
        year = col_number(), population = col_number()))
complete_pops <- complete_pops %>%
  rename("pop" = "population") %>%
  mutate(VARNAME_2 = str_to_title(VARNAME_2))

```

## Continue Cleaning
```{r}

# fix the NA for Tan An, for some reason was given the name Ben Luc when coded without symbols, also fix Cai Lay and An Phu
updated_MDR_Dengue_Cases$no_symbols_name[updated_MDR_Dengue_Cases$district == "tân an" & updated_MDR_Dengue_Cases$no_symbols_name ==  "ben luc"] <- "tan an"
updated_MDR_Dengue_Cases$no_symbols_name[updated_MDR_Dengue_Cases$district == "tân an" & updated_MDR_Dengue_Cases$no_symbols_name ==  "ben luc"] <- "tan an"
# need to run again to fix Tan An
updated_MDR_Dengue_Cases <- updated_MDR_Dengue_Cases %>%
  mutate(district_type = case_when(
    no_symbols_name %in% City ~ "city", 
    no_symbols_name %in% Town ~ "town", 
    no_symbols_name %in% Urban_District ~ "urban district", 
    no_symbols_name %in% District ~ "district"
  ))

updated_MDR_Dengue_Cases_1 <- updated_MDR_Dengue_Cases %>%
  mutate(no_symbols_name_join = no_symbols_name, 
         no_symbols_name_join = tolower(no_symbols_name_join))



```

## Get Province Map
```{r}
MDR_2 <- st_read(dsn = '/Users/zoe/Documents/Yale EMD/Rotation 1/Dengue - Vietnam/dengue_exploration/Data/MDR Updated map/VMD_districts_2022.shp')

MDR_2 %>%
  mutate(MDR_2, VARNAME_2 = as.character(VARNAME_2)) %>%
  mutate(MDR_2, ENGTYPE_2 = as.character(ENGTYPE_2)) 

```

## Clean & account for districts/provinces that have split off
```{r}

# re-capitalize district names and types to match with the shapefile
updated_MDR_Dengue_Cases <- updated_MDR_Dengue_Cases %>%
  mutate(no_symbols_name = str_to_title(no_symbols_name), 
         district_type = str_to_title(district_type))
updated_monthly_DHF_incidence <- updated_monthly_DHF_incidence %>%
  mutate(no_symbols_name = str_to_title(no_symbols_name), 
         district_type = str_to_title(district_type))

### check why map is missing districts

#varnames <- rep(MDR_3$VARNAME_2, 19)
#casenames <- updated_cases$no_symbols_name

#check_names <- data.frame(c(unique(varnames), rep(NA, 3)), unique(casenames))
# varnames is missing Can Tho (code: 28-31, 33, 35 --> 2001-2003), Mo Cay (becomes Mo Cay Bac and Mo Cay Nam in 2009), Duyen Hai (122 and 123, 123 should be a town), Cai Lay (108 and 109, just 109 after 2015 when should be a town, different dengue numbers), Long My (60 and 61, 61 should be a town), Cao Lanh (46 and 47, 47 should be a city), Hong Ngu (49 and 50, 50 should be a city)

### fix mapping problem by setting incidence equal in Mo Cay Bac and Mo Cay Nam 2001 - 2008
updated_MDR_Dengue_Cases_map_1 <- updated_MDR_Dengue_Cases %>%
  filter(no_symbols_name == "Mo Cay") %>%
  mutate(no_symbols_name = str_replace(no_symbols_name, "Mo Cay", "Mo Cay Bac"))
updated_MDR_Dengue_Cases_map_2 <- updated_MDR_Dengue_Cases %>%
  filter(no_symbols_name == "Mo Cay") %>%
  mutate(no_symbols_name = str_replace(no_symbols_name, "Mo Cay", "Mo Cay Nam"))
updated_MDR_Dengue_Cases_map <- updated_MDR_Dengue_Cases %>%
  rbind(updated_MDR_Dengue_Cases_map_1, updated_MDR_Dengue_Cases_map_2)
updated_MDR_Dengue_Cases_map <- updated_MDR_Dengue_Cases_map %>%
  dplyr::filter(updated_MDR_Dengue_Cases_map$no_symbols_name != "Mo Cay")

### fix map problem by setting incidence equal in Binh Thuy (28), Co Do (29), Cai Rang (30), Ninh Kieu (31), Phong Dien (33), Thoi Lai (35) --> 2001-2003
updated_MDR_Dengue_Cases_map_3 <- updated_MDR_Dengue_Cases %>%
  filter(no_symbols_name == "Can Tho") %>%
  mutate(no_symbols_name = str_replace(no_symbols_name, "Can Tho", "Binh Thuy"))
updated_MDR_Dengue_Cases_map_4 <- updated_MDR_Dengue_Cases %>%
  filter(no_symbols_name == "Can Tho") %>%
  mutate(no_symbols_name = str_replace(no_symbols_name, "Can Tho", "Co Do"))
updated_MDR_Dengue_Cases_map_5 <- updated_MDR_Dengue_Cases %>%
  filter(no_symbols_name == "Can Tho") %>%
  mutate(no_symbols_name = str_replace(no_symbols_name, "Can Tho", "Cai Rang"))
updated_MDR_Dengue_Cases_map_6 <- updated_MDR_Dengue_Cases %>%
  filter(no_symbols_name == "Can Tho") %>%
  mutate(no_symbols_name = str_replace(no_symbols_name, "Can Tho", "Ninh Kieu"))
updated_MDR_Dengue_Cases_map_7 <- updated_MDR_Dengue_Cases %>%
  filter(no_symbols_name == "Can Tho") %>%
  mutate(no_symbols_name = str_replace(no_symbols_name, "Can Tho", "Phong Dien"))
updated_MDR_Dengue_Cases_map_8 <- updated_MDR_Dengue_Cases %>%
  filter(no_symbols_name == "Can Tho") %>%
  mutate(no_symbols_name = str_replace(no_symbols_name, "Can Tho", "Thoi Lai"))
updated_MDR_Dengue_Cases_map <- updated_MDR_Dengue_Cases_map %>%
  rbind(updated_MDR_Dengue_Cases_map_3, updated_MDR_Dengue_Cases_map_4, updated_MDR_Dengue_Cases_map_5, updated_MDR_Dengue_Cases_map_6, updated_MDR_Dengue_Cases_map_7, updated_MDR_Dengue_Cases_map_8)
updated_MDR_Dengue_Cases_map <- updated_MDR_Dengue_Cases_map %>%
  dplyr::filter(updated_MDR_Dengue_Cases_map$no_symbols_name != "Can Tho")

# fix mapping problem --> Duyen Hai (122 and 123, 123 should be a town)
updated_MDR_Dengue_Cases_map_9 <- updated_MDR_Dengue_Cases_map %>%
  filter(`updated_district_code.y` == 123) %>%
  mutate(district_type = case_when(
    `updated_district_code.y` == 123 ~ "Town"))
updated_MDR_Dengue_Cases_map <- updated_MDR_Dengue_Cases_map %>%
  rbind(updated_MDR_Dengue_Cases_map_9)


# fix mapping problem --> Cai Lay (108 and 109, just 109 after 2015 when should be a town, different dengue numbers)
updated_MDR_Dengue_Cases_map_10 <- updated_MDR_Dengue_Cases_map %>%
  filter(`updated_district_code.y` == 109) %>%
  mutate(district_type = case_when(
    `updated_district_code.y` == 109 ~ `district_type` == "Town"))
updated_MDR_Dengue_Cases_map <- updated_MDR_Dengue_Cases_map %>%
  rbind(updated_MDR_Dengue_Cases_map_10)

# fix mapping problem --> Long My (60 and 61, 61 should be a town)
updated_MDR_Dengue_Cases_map_11 <- updated_MDR_Dengue_Cases_map %>%
   filter(`updated_district_code.y` == 61) %>%
   mutate(district_type = case_when(
    `updated_district_code.y` == 61 ~ "Town")) %>%
  mutate(district_type = case_when(
    `updated_district_code.x` == 61 ~ "Town"))
updated_MDR_Dengue_Cases_map <- updated_MDR_Dengue_Cases_map %>%
  rbind(updated_MDR_Dengue_Cases_map_11) 

# fix mapping problem --> Cao Lanh (46 and 47, 47 should be a city)
updated_MDR_Dengue_Cases_map_12 <- updated_MDR_Dengue_Cases_map %>%
  filter(`updated_district_code.y` == 47) %>%
  mutate(district_type = case_when(
    `updated_district_code.y` == 47 ~ "City"))
updated_MDR_Dengue_Cases_map <- updated_MDR_Dengue_Cases_map %>%
  rbind(updated_MDR_Dengue_Cases_map_12)

# fix mapping problem --> Hong Ngu (49 and 50, 50 should be a city)
updated_MDR_Dengue_Cases_map_13 <- updated_MDR_Dengue_Cases_map %>%
  filter(`updated_district_code.y` == 50) %>%
  mutate(district_type = case_when(
    `updated_district_code.y` == 50 ~ "City"))
updated_MDR_Dengue_Cases_map <- updated_MDR_Dengue_Cases_map %>%
  rbind(updated_MDR_Dengue_Cases_map_13)

# Cai Lay 109 should be a town
updated_MDR_Dengue_Cases_map_14 <- updated_MDR_Dengue_Cases_map %>%
  filter(`updated_district_code.y` == 109) %>%
  mutate(district_type = case_when(
    `updated_district_code.y` == 109 ~ "Town"))
updated_MDR_Dengue_Cases_map <- updated_MDR_Dengue_Cases_map %>%
  rbind(updated_MDR_Dengue_Cases_map_14)

### make provinces match the ones in the shapefile
updated_MDR_Dengue_Cases_map <- updated_MDR_Dengue_Cases_map %>%
  mutate(province = str_replace(province, "An Giang", "An Giang"), 
         province = str_replace(province, "Bac Lieu", "Bạc Liêu"), 
         province = str_replace(province, "Ben Tre" , "Bến Tre"), 
         province = str_replace(province, "Ca Mau" , "Cà Mau"),
         province = str_replace(province, "Can Tho" , "Cần Thơ"),
         province = str_replace(province, "Dong Thap" , "Đồng Tháp"),
         province = str_replace(province, "Hau Giang" , "Hậu Giang"),
         province = str_replace(province, "Kien Giang" , "Kiên Giang"),
         province = str_replace(province, "Long An" , "Long An"),
         province = str_replace(province, "Soc Trang" , "Sóc Trăng"),
         province = str_replace(province, "Tien Giang" , "Tiền Giang"),
         province = str_replace(province, "Tra Vinh" , "Trà Vinh"),
         province = str_replace(province, "Vinh Long" , "Vĩnh Long"))



```

## Combine Map and Cases
```{r}
# join by province, district, and district type
MDR_Dengue_map_pre <- left_join(MDR_2, updated_MDR_Dengue_Cases_map, by = c("NAME_1" = "province", "VARNAME_2" = "no_symbols_name", "ENGTYPE_2" = "district_type"))


## has not been updated
MDR_monthly_dengue_map <- left_join(MDR_2, updated_monthly_DHF_incidence, by = c("VARNAME_2" = "no_symbols_name", "ENGTYPE_2" = "district_type"))

```

## Create Incidence 
```{r}
MDR_Dengue_map_1 <- left_join(MDR_Dengue_map_pre, complete_pops, by = c("VARNAME_2" = "VARNAME_2", "NAME_1" = "NAME_1", "ENGTYPE_2" = "ENGTYPE_2", "year" = "year")) 

MDR_Dengue_map_1 <- MDR_Dengue_map_1 %>%
  dplyr::select(!(GID_0)) %>%
  dplyr::select(!(NAME_0)) %>%
  dplyr::select(!(GID_1)) %>%
  dplyr::select(!(GID_2)) %>%
  dplyr::select(!(HASC_2)) %>%
  dplyr::select(!(stt)) %>%
  dplyr::select(!(NL_NAME_1)) %>%
  dplyr::select(!(NL_NAME_2)) %>%
  dplyr::select(!(CC_2)) %>%
  dplyr::select(!(district_code)) %>%
  dplyr::select(!(updated_district_code.x)) %>%
  dplyr::select(!(population)) %>%
  dplyr::select(!(`population(thousand)`))

# create incidence per 100,000 people column
MDR_Dengue_map <- MDR_Dengue_map_1 %>%
  mutate(DHF_incidence = 100000*(total_DHF_cases/pop)) %>%
  mutate(log_DHF_incidence = log(DHF_incidence)) %>%
  mutate(m1 = as.numeric(m1), 
         m2 = as.numeric(m2),
         m3 = as.numeric(m3), 
         m4 = as.numeric(m4), 
         m5 = as.numeric(m5), 
         m6 = as.numeric(m6))

MDR_Dengue_map <- MDR_Dengue_map %>%
  mutate(log_DHF_incidence = ifelse(log_DHF_incidence == -Inf, 0, log_DHF_incidence))

# create a monthly incidence df
updated_monthly_DHF_incidence <- MDR_Dengue_map %>%
  pivot_longer(cols = starts_with("m"), 
               values_to = "m_DHF_cases", 
               names_to = "monthvar") %>%
  mutate(m_DHF_incidence = 100000*(m_DHF_cases/pop), 
         log_m_DHF_incidence = log(m_DHF_incidence))
```


## An Giang Province
```{r}

an_giang_cases_by_year <- updated_MDR_Dengue_Cases %>%
  filter(province == "An Giang") %>%
  group_by(year) %>%
  summarize(year_total = sum(total_DHF_cases))

an_giang_cases_by_year %>%
  ggplot(aes(x = year, y = year_total))+
  geom_bar(stat = "identity")+
  geom_jitter(width = 0.15)+
  theme(axis.text.x = element_text(angle = 90))+
  labs(title = "An Giang Province DHF Cases 2001 - 2018", x = "Year", y = "Total DHF Cases")

# look at time series
an_giang_cases <- updated_MDR_Dengue_Cases %>%
  filter(province == "An Giang")

ggplot(an_giang_cases, aes(x = year, y = total_DHF_cases, group=district, col = district))+
  geom_line()+
  theme_classic()
```

## Bac Lieu Province
```{r}

bac_lieu_cases_by_year <- updated_MDR_Dengue_Cases %>%
  filter(province == "Bac Lieu") %>%
  group_by(year) %>%
  summarize(year_total = sum(total_DHF_cases))
  
bac_lieu_cases_by_year %>%
  ggplot(aes(x = year, y = year_total))+
  geom_bar(stat = "identity")+
  geom_jitter(width = 0.15)+
  theme(axis.text.x = element_text(angle = 90))+
  labs(title = "Bac Lieu Province DHF Cases 2001 - 2018", x = "Year", y = "Total DHF Cases")
  
# look at time series
bac_lieu_cases <- updated_MDR_Dengue_Cases %>%
  filter(province == "Bac Lieu")

ggplot(bac_lieu_cases, aes(x = year, y = total_DHF_cases, group=district, col = district))+
  geom_line()+
  theme_classic()

```
## Ben Tre Province
```{r}

ben_tre_cases_by_year <- updated_MDR_Dengue_Cases %>%
  filter(province == "Ben Tre") %>%
  group_by(year) %>%
  summarize(year_total = sum(total_DHF_cases))
  
ben_tre_cases_by_year %>%
  ggplot(aes(x = year, y = year_total))+
  geom_bar(stat = "identity")+
  geom_jitter(width = 0.15)+
  theme(axis.text.x = element_text(angle = 90))+
  labs(title = "Ben Tre Province DHF Cases 2001 - 2018", x = "Year", y = "Total DHF Cases")

# look at time series
ben_tre_cases <- updated_MDR_Dengue_Cases %>%
  filter(province == "Ben Tre")

ggplot(ben_tre_cases, aes(x = year, y = total_DHF_cases, group=district, col = district))+
  geom_line()+
  theme_classic()

```
## Ca Mau Province
```{r}

ca_mau_cases_by_year <- updated_MDR_Dengue_Cases %>%
  filter(province == "Ca Mau") %>%
  group_by(year) %>%
  summarize(year_total = sum(total_DHF_cases))
  
ca_mau_cases_by_year %>%
  ggplot(aes(x = year, y = year_total))+
  geom_bar(stat = "identity")+
  geom_jitter(width = 0.15)+
  theme(axis.text.x = element_text(angle = 90))+
  labs(title = "Ca Mau Province DHF Cases 2001 - 2018", x = "Year", y = "Total DHF Cases")

# look at time series
ca_mau_cases <- updated_MDR_Dengue_Cases %>%
  filter(province == "Ca Mau")

ggplot(ca_mau_cases, aes(x = year, y = total_DHF_cases, group=district, col = district))+
  geom_line()+
  theme_classic()
```

## Can Tho Province
```{r}

can_tho_cases_by_year <- updated_MDR_Dengue_Cases %>%
  filter(province == "Can Tho") %>%
  group_by(year) %>%
  summarize(year_total = sum(total_DHF_cases))
  
can_tho_cases_by_year %>%
  ggplot(aes(x = year, y = year_total))+
  geom_bar(stat = "identity")+
  geom_jitter(width = 0.15)+
  theme(axis.text.x = element_text(angle = 90))+
  labs(title = "Can Tho Province DHF Cases 2001 - 2018", x = "Year", y = "Total DHF Cases")

# look at time series
can_tho_cases <- updated_MDR_Dengue_Cases %>%
  filter(province == "Can Tho")

ggplot(can_tho_cases, aes(x = year, y = total_DHF_cases, group=district, col = district))+
  geom_line()+
  theme_classic()

#### weird jump just before 2005
```

## Dong Thap Province
```{r}

dong_thap_cases_by_year <- updated_MDR_Dengue_Cases %>%
  filter(province == "Dong Thap") %>%
  group_by(year) %>%
  summarize(year_total = sum(total_DHF_cases))
  
dong_thap_cases_by_year %>%
  ggplot(aes(x = year, y = year_total))+
  geom_bar(stat = "identity")+
  geom_jitter(width = 0.15)+
  theme(axis.text.x = element_text(angle = 90))+
  labs(title = "Dong Thap Province DHF Cases 2001 - 2018", x = "Year", y = "Total DHF Cases")

# look at time series
dong_thap_cases <- updated_MDR_Dengue_Cases %>%
  filter(province == "Dong Thap")

ggplot(dong_thap_cases, aes(x = year, y = total_DHF_cases, group=district, col = district))+
  geom_line()+
  theme_classic()

```

## Hau Giang Province
```{r}

hau_giang_cases_by_year <- updated_MDR_Dengue_Cases %>%
  filter(province == "Hau Giang") %>%
  group_by(year) %>%
  summarize(year_total = sum(total_DHF_cases))
  
hau_giang_cases_by_year %>%
  ggplot(aes(x = year, y = year_total))+
  geom_bar(stat = "identity")+
  geom_jitter(width = 0.15)+
  theme(axis.text.x = element_text(angle = 90))+
  labs(title = "Hau Giang Province DHF Cases 2004 - 2018", x = "Year", y = "Total DHF Cases")

# look at time series
hau_giang_cases <- updated_MDR_Dengue_Cases %>%
  filter(province == "Hau Giang")

ggplot(hau_giang_cases, aes(x = year, y = total_DHF_cases, group=district, col = district))+
  geom_line()+
  theme_classic()
```

## Kien Giang Province
```{r}

kien_giang_cases_by_year <- updated_MDR_Dengue_Cases %>%
  filter(province == "Kien Giang") %>%
  group_by(year) %>%
  summarize(year_total = sum(total_DHF_cases))
  
kien_giang_cases_by_year %>%
  ggplot(aes(x = year, y = year_total))+
  geom_bar(stat = "identity")+
  geom_jitter(width = 0.15)+
  theme(axis.text.x = element_text(angle = 90))+
  labs(title = "Kien Giang Province DHF Cases 2001 - 2018", x = "Year", y = "Total DHF Cases")

# look at time series
kien_giang_cases <- updated_MDR_Dengue_Cases %>%
  filter(province == "Kien Giang")

ggplot(kien_giang_cases, aes(x = year, y = total_DHF_cases, group=district, col = district))+
  geom_line()+
  theme_classic()

```

## Long An Province
```{r}

long_an_cases_by_year <- updated_MDR_Dengue_Cases %>%
  filter(province == "Long An") %>%
  group_by(year) %>%
  summarize(year_total = sum(total_DHF_cases))
  
long_an_cases_by_year %>%
  ggplot(aes(x = year, y = year_total))+
  geom_bar(stat = "identity")+
  geom_jitter(width = 0.15)+
  theme(axis.text.x = element_text(angle = 90))+
  labs(title = "Long An Province DHF Cases 2001 - 2018", x = "Year", y = "Total DHF Cases")

# look at time series
long_an_cases <- updated_MDR_Dengue_Cases %>%
  filter(province == "Long An")

ggplot(long_an_cases, aes(x = year, y = total_DHF_cases, group=district, col = district))+
  geom_line()+
  theme_classic()

```

## Soc Trang Province
```{r}

soc_trang_cases_by_year <- updated_MDR_Dengue_Cases %>%
  filter(province == "Soc Trang") %>%
  group_by(year) %>%
  summarize(year_total = sum(total_DHF_cases))
  
soc_trang_cases_by_year %>%
  ggplot(aes(x = year, y = year_total))+
  geom_bar(stat = "identity")+
  geom_jitter(width = 0.15)+
  theme(axis.text.x = element_text(angle = 90))+
  labs(title = "Soc Trang Province DHF Cases 2001 - 2018", x = "Year", y = "Total DHF Cases")

# look at time series
soc_trang_cases <- updated_MDR_Dengue_Cases %>%
  filter(province == "Soc Trang")

ggplot(soc_trang_cases, aes(x = year, y = total_DHF_cases, group=district, col = district))+
  geom_line()+
  theme_classic()

```

## Tien Giang Province
```{r}

tien_giang_cases_by_year <- updated_MDR_Dengue_Cases%>%
  filter(province == "Tien Giang") %>%
  group_by(year) %>%
  summarize(year_total = sum(total_DHF_cases))
  
tien_giang_cases_by_year %>%
  ggplot(aes(x = year, y = year_total))+
  geom_bar(stat = "identity")+
  geom_jitter(width = 0.15)+
  theme(axis.text.x = element_text(angle = 90))+
  labs(title = "Tien Giang Province DHF Cases 2001 - 2018", x = "Year", y = "Total DHF Cases")

# look at time series
tien_giang_cases <- updated_MDR_Dengue_Cases %>%
  filter(province == "Tien Giang")

ggplot(tien_giang_cases, aes(x = year, y = total_DHF_cases, group=district, col = district))+
  geom_line()+
  theme_classic()

```

## Tra Vinh Province
```{r}

tra_vinh_cases_by_year <- updated_MDR_Dengue_Cases %>%
  filter(province == "Tra Vinh") %>%
  group_by(year) %>%
  summarize(year_total = sum(total_DHF_cases))
  
tra_vinh_cases_by_year %>%
  ggplot(aes(x = year, y = year_total))+
  geom_bar(stat = "identity")+
  geom_jitter(width = 0.15)+
  theme(axis.text.x = element_text(angle = 90))+
  labs(title = "Tra Vinh Province DHF Cases 2001 - 2018", x = "Year", y = "Total DHF Cases")

# look at time series
tra_vinh_cases <- updated_MDR_Dengue_Cases %>%
  filter(province == "Tra Vinh")

ggplot(tra_vinh_cases, aes(x = year, y = total_DHF_cases, group=district, col = district))+
  geom_line()+
  theme_classic()


```

## Vinh Long Province
```{r}

vinh_long_cases_by_year <- updated_MDR_Dengue_Cases %>%
  filter(province == "Vinh Long") %>%
  group_by(year) %>%
  summarize(year_total = sum(total_DHF_cases))
  
vinh_long_cases_by_year %>%
  ggplot(aes(x = year, y = year_total))+
  geom_bar(stat = "identity")+
  geom_jitter(width = 0.15)+
  theme(axis.text.x = element_text(angle = 90))+
  labs(title = "Vinh Long Province DHF Cases 2001 - 2018", x = "Year", y = "Total DHF Cases")

# look at time series
vinh_long_cases <- updated_MDR_Dengue_Cases %>%
  filter(province == "Vinh Long")

ggplot(vinh_long_cases, aes(x = year, y = total_DHF_cases, group=district, col = district))+
  geom_line()+
  theme_classic()

```

## Basic Time Series for All Districts/Provinces
```{r}

# total DHF cases
ggplot(updated_MDR_Dengue_Cases, aes(x = year, y = total_DHF_cases, group=district, col = province))+
  geom_line()+
  theme_classic()+
  labs(title = "MDR Province Total DHF Cases 2001-2018", x = "Year", y = "Total DHF Cases")

ggplot(updated_MDR_Dengue_Cases, aes(x = year, y = total_DHF_cases, group=district))+
  geom_line()+
  facet_wrap(~province, scales = 'free_y')+
  theme_classic() +
  labs(title = "MDR Province Total DHF Cases 2001-2018", subtitle = "Grouped by Province", x = "Year", y = "Total DHF Cases")

ggplot(updated_MDR_Dengue_Cases, aes(x = year, y = total_DHF_cases, group=district, col = province))+
  geom_line()+
  facet_wrap(~factor(district_type), scales = 'free_y')+
  theme_classic()+
  labs(title = "MDR Province Total DHF Cases 2001-2018", subtitle = "Grouped by District Type", x = "Year", y = "Total DHF Cases")

ggplot(updated_monthly_DHF_incidence, aes(x = monthvar, y = m_DHF_cases, group=district, col = province))+
  geom_line()+
  facet_wrap(~factor(year), scales = 'free_y')+
  theme_classic()+
  theme(axis.text.x = element_text(angle = 90, size = 5))+ 
  labs(title = "MDR Province Total DHF Cases", subtitle = "Grouped by Year", x = "Month", y = "Total DHF Cases")

ggplot(updated_monthly_DHF_incidence, aes(x = year, y = m_DHF_cases, group=district, col = province))+
  geom_line()+
  facet_wrap(~factor(monthvar), scales = 'free_y')+
  theme_classic()+
  theme(axis.text.x = element_text(angle = 90, size = 5))+ 
  labs(title = "MDR Province Total DHF Cases", subtitle = "Grouped by Month", x = "Year", y = "Total DHF Cases")

# DHF incidence 
ggplot(updated_MDR_Dengue_Cases, aes(x = year, y = DHF_incidence, group=district, col = province))+
  geom_line()+
  theme_classic()+
  labs(title = "MDR Province DHF Incidence 2001-2018", subtitle = "Using current population estimates", x = "Year", y = "DHF Incidence per 100,000 People")

ggplot(updated_MDR_Dengue_Cases, aes(x = year, y = DHF_incidence, group=district))+
  geom_line()+
  facet_wrap(~province, scales = 'free_y')+
  theme_classic() +
  labs(title = "MDR Province DHF Incidence 2001-2018", subtitle = "Grouped by province, using current population estimates", x = "Year", y = "DHF Incidence per 100,000 People")

ggplot(updated_MDR_Dengue_Cases, aes(x = year, y = DHF_incidence, group=district, col = province))+
  geom_line()+
  facet_wrap(~factor(district_type), scales = 'free_y')+
  theme_classic()+
  labs(title = "MDR Province DHF Incidence 2001-2018", subtitle = "Grouped by district type, using current population estimates", x = "Year", y = "DHF Incidence per 100,000 People")

ggplot(updated_monthly_DHF_incidence, aes(x = monthvar, y = m_DHF_incidence, group=district, col = province))+
  geom_line()+
  facet_wrap(~factor(year), scales = 'free_y')+
  theme_classic()+
  theme(axis.text.x = element_text(angle = 90, size = 5))+ 
  labs(title = "MDR Province DHF Incidence", subtitle = "Grouped by Year", x = "Month", y = "DHF Incidence per 100,000 People")

ggplot(updated_monthly_DHF_incidence, aes(x = year, y = m_DHF_incidence, group=district, col = province))+
  geom_line()+
  facet_wrap(~factor(monthvar), scales = 'free_y')+
  theme_classic()+
  theme(axis.text.x = element_text(angle = 90, size = 5))+ 
  labs(title = "MDR Province DHF Incidence", subtitle = "Grouped by Month", x = "Year", y = "DHF Incidence per 100,000 People")

```



## Fix Mapping DF for mapping
```{r}

MDR_Dengue_map_1 <- fortify(MDR_Dengue_map) %>%
  dplyr::arrange(NAME_1, VARNAME_2, ENGTYPE_2, year) %>%
  group_by(NAME_1, VARNAME_2, ENGTYPE_2, year) %>%
  distinct()

```



## Map DHF Incidence Data 
```{r}
library(plotly)
# total case map
g <- ggplot(data = MDR_Dengue_map)+
  geom_sf(aes(fill = total_DHF_cases, names = VARNAME_2), color = "black", size = 0.25)+
  scale_fill_distiller(name = "DHF Cases", 
                       palette = "Blues", 
                       direction = 1,
                       breaks = pretty_breaks())+
   theme(axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks = element_blank(),
        panel.background = element_rect(fill = "white", color = NA)) +
  labs(title = "Total DHF Cases in MDR 2001-2018")
 #ggplotly(g)
g

# case map with district population included 
st_geometry_type(MDR_Dengue_map, by_geometry=FALSE)
sdcntrd <- st_centroid(MDR_Dengue_map)
st_geometry_type(MDR_Dengue_map, by_geometry = FALSE)

ggplot()+
  geom_sf(data = MDR_Dengue_map_1, aes(fill = total_DHF_cases), color = "black", size = 0.25)+
  scale_fill_distiller(name = "DHF Cases", 
                       palette = "Blues", 
                       direction = 1,
                       breaks = pretty_breaks())+
  geom_sf(data = sdcntrd, aes(size = pop), color = "black", show.legend = "point")+
  scale_size(name = "Population")+
   theme(axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks = element_blank(),
        panel.background = element_rect(fill = "white", color = NA)) +
  labs(title = "Total DHF Cases in MDR 2001-2018")

# incidence map
ggplot(data = MDR_Dengue_map)+
  geom_sf(aes(fill = DHF_incidence), lwd = 0, size = 0.25)+
  scale_fill_distiller(name = "DHF Incidence per 100,000 people", 
                       palette = "Reds", 
                       direction = 1,
                       breaks = pretty_breaks())+
   theme(axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks = element_blank(),
        panel.background = element_rect(fill = "white", color = NA)) +
  labs(title = "DHF Incidence in MDR 2001-2018")

ggplot(data = MDR_Dengue_map_1)+
  geom_sf(aes(fill = log(DHF_incidence)), lwd = 0, size = 0.25)+
  scale_fill_distiller(name = "Log DHF Incidence per 100,000 people", 
                       palette = "Reds", 
                       direction = 1,
                       breaks = pretty_breaks())+
   theme(axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks = element_blank(),
        panel.background = element_rect(fill = "white", color = NA)) +
  labs(title = "DHF Incidence in MDR 2001-2018")

#incidence maps by year
ggplot(data = MDR_Dengue_map_1)+
  geom_sf(aes(fill = log(DHF_incidence)), lwd = 0, size = 0.25)+
  facet_wrap(~factor(year))+
  scale_fill_distiller(name = "Log DHF Incidence per 100,000 people", 
                       palette = "Reds", 
                       direction = 1,
                       breaks = pretty_breaks())+
   theme(axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks = element_blank(),
        panel.background = element_rect(fill = "white", color = NA)) +
  labs(title = "DHF Incidence in MDR 2001-2018", subtitle = "Grouped by year")

#incidence maps by month
ggplot(data = MDR_monthly_dengue_map)+
  geom_sf(aes(fill = log(m_DHF_incidence)), lwd = 0, size = 0.25)+
  facet_wrap(~factor(monthvar))+
  scale_fill_distiller(name = "Log DHF Incidence per 100,000 people", 
                       palette = "Reds", 
                       direction = 1,
                       breaks = pretty_breaks())+
   theme(axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks = element_blank(),
        panel.background = element_rect(fill = "white", color = NA)) +
  labs(title = "DHF Incidence in MDR 2001-2018", subtitle = "Grouped by month")


# average of each month by district 
MDR_monthly_average_map <- MDR_Dengue_map %>%
  group_by(factor(VARNAME_2)) %>%
  summarize(m1_avg = mean(m1),
            m2_avg = mean(m2),
            m3_avg = mean(m3),
            m4_avg = mean(m4),
            m5_avg = mean(m5),
            m6_avg = mean(m6),
            m7_avg = mean(m7),
            m8_avg = mean(m8),
            m9_avg = mean(m9),
            m10_avg = mean(m10),
            m11_avg = mean(m11),
            m12_avg = mean(m12))
  
MDR_monthly_average_map <- MDR_monthly_average_map %>%
  pivot_longer(cols = starts_with("m"), 
               values_to = "m_avg_DHF_cases", 
               names_to = "month")

  ggplot(data = MDR_monthly_average_map)+
  geom_sf(aes(fill = m_avg_DHF_cases, na.rm = T), lwd = 0, size = 0.25)+
  facet_wrap(~factor(month))+
  scale_fill_distiller(name = "DHF Cases", 
                       palette = "Reds", 
                       direction = 1,
                       breaks = pretty_breaks())+
   theme(axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks = element_blank(),
        panel.background = element_rect(fill = "white", color = NA)) +
  labs(title = "Average Monthly DHF Cases 2001 - 2018", subtitle = "Grouped by month")
  
  
```


## Heat Map
```{r}


MDR_dengue_heatmap <- MDR_Dengue_map_1 %>%
  group_by(VARNAME_2) %>%
  mutate(scaled_cases = total_DHF_cases/max(total_DHF_cases))

MDR_monthly_dengue_heatmap <- MDR_monthly_dengue_map %>%
 group_by(VARNAME_2) %>%
  group_by(monthvar) %>%
  mutate(scaled_cases = m_DHF_cases/max(m_DHF_cases))

# by each district
ggplot(MDR_dengue_heatmap, aes(year, VARNAME_2, fill = scaled_cases))+
  geom_tile()+ 
  scale_fill_distiller(name = "Scaled DHF Cases", 
                       palette = "Blues", 
                       direction = 1,
                       breaks = pretty_breaks())+
  theme(axis.text.y = element_text(size = 3))+
  labs(x = "Year", y = "District", title ="Heatmap of DHF Cases in MDR 2001 - 2018")


# by each province
ggplot(MDR_dengue_heatmap, aes(year, NAME_1, fill = scaled_cases))+
  geom_tile() +
  scale_fill_distiller(name = "Scaled DHF Cases", 
                       palette = "Blues", 
                       direction = 1,
                       breaks = pretty_breaks())+
  labs(x = "Year", y = "Province", title = "Heatmap of MDR DHF Cases in MDR 2001 - 2018")

#by each month
ggplot(MDR_monthly_dengue_heatmap, aes(year, monthvar, fill = scaled_cases))+
  geom_tile() +
  labs(x = "Year", y = "Month", title = "Heatmap of MDR DHF Cases")

```

