---
title: "Model Building"
author: "Elisabeth Nelson"
date: "2022-10-20"
output: pdf_document
---
## Load Packages
```{r}
#library(rjags)
library(MCMCpack)
library(stats)
library(CARBayes) # <- this is the package for the model
library(rgdal) # <- for working with shapefiles
library(RColorBrewer) 
library(ggplot2)
library(rgeos) # <- also for working with shapefiles
library(maptools) # shapefiles again
library(spdep) # still shapefiles 
library(ggmap) # plotting shapefiles 
library(sf)
library(dplyr)
library(tidyverse)
library(CARBayesST)

```


## Set-up for Carbays
```{r}
# Number of spatial units (e.g., districts in MDR)
n.district <- 134

# Create an n x􏰀 n identity matrix
Id.mat <- diag(n.district)

MDR_2 <- st_read(dsn = '/Users/zoe/Documents/Yale EMD/Rotation 1/Dengue - Vietnam/dengue_exploration/Data/MDR Updated map/VMD_districts_2022.shp')

```

## Neighbor Matrix
```{r}
# Neighbors can either be Queen (any zip that touches another zip - even at 
# corners) or Rook neighbors (share full sides -- not corners)

# The snap dist is governing the NA in neighbor matrix
# Can remove SNAP statement; if Queen = F then single point touch not allowed 
# and Rook distance used.

# When data is missing -- there's more non-zero links involved 
# Try to make snap distance as small as possible while making sure all states 
# have at least one neighbor
neighb <- poly2nb(MDR_2, queen = T, snap = sqrt(0.001))

# Check average number of links -- increase snap distance if it says there is 
# a zip with zero neighbors
neighb # Avg Number of links = 5.19403 
  # regions 74 nand 76 have no links --> islands Kien Hai/District + Phu Quoc/City

# Make neighbor matrix 
# if zero.policy FALSE stop with error for any empty neighbour sets, 
# if TRUE permit the weights list to be formed with zero-length weights vectors
# B is the basic binary coding, W is row standardised (both are commonly used)
neighb.mat <- nb2mat(neighb, zero.policy = T, style = 'B')

    # the island will not have neighbors, so need to exclude it at first
neighb.mat_1 <- neighb.mat[-74, -74]
neighb.mat_2 <- neighb.mat_1[-75, -75]



# For Leroux model, neighbor matrix needs to have 0 if not neighboring, -1 if 
# neighbor, and # of neighbors for each state on diagonal), so...
w.mat <- ifelse(neighb.mat==1,-1,neighb.mat)
num.neigh <- colSums(neighb.mat) # Count the number of neighbors for each state
for (i in 1:nrow(w.mat)) { # Put the number of neighbors on diagonal
  for (j in 1:ncol(w.mat)) {
    if (i==j) {
      w.mat[i,j] <- num.neigh[i]
    }
  }
}
# w.mat <- array(NA, dim=c(n.stzipcode,n.zipcode,n.countries))
# w.mat[,,1] <- ifelse(neighb.mat==1,-1,neighb.mat) # Replace 1 with -1
# num.neigh <- colSums(neighb.mat) # Count the number of neighbors for each state
# for (i in 1:nrow(w.mat[,,1])) { # Put the number of neighbors on diagonal
#   for (j in 1:ncol(w.mat[,,1])) {
#     if (i==j) {
#       w.mat[i,j,1] <- num.neigh[i]
#     }
#   }
# }


# could eventually define neighbors for the islands based on spatial adjacency/transport
# could also separately analyze them once we have a model and see if it predcits well in them independently

```

## Set up Data Frame for Model Training
```{r}

# create data frame
base_model_2001_2006 <- fortify(MDR_Dengue_map) %>%
  dplyr::select(NAME_1, VARNAME_2, ENGTYPE_2, district, total_DHF_cases, year, m1, m2, m3, m4, m5, m6, m7, m8, m9, m10, m11, m12, updated_district_code.y, pop, DHF_incidence, log_DHF_incidence, geometry) 

base_model_2001_2006 <- base_model_2001_2006 %>% 
  filter(between(year, 2001, 2006))

# fix districts that are missing years 
colnames(base_model_2001_2006)

geom_1 <- MDR_Dengue_map[1033, ]$geometry

correction_df <- data.frame(NAME_1 = c(rep("Hậu Giang", 3)),
                            VARNAME_2 = c(rep("Chau Thanh A", 3)),
                            ENGTYPE_2 = c(rep("District", 3)),
                            district = c(rep("châu thành a", 3)),
                            total_DHF_cases = c(rep(NA, 3)),
                            year = c(2001, 2002, 2003),
                            m1 = c(rep(NA, 3)), 
                            m2 = c(rep(NA, 3)), 
                            m3 = c(rep(NA, 3)), 
                            m4 = c(rep(NA, 3)), 
                            m5 = c(rep(NA, 3)), 
                            m6 = c(rep(NA, 3)),
                            m7 = c(rep(NA, 3)),
                            m8 = c(rep(NA, 3)),
                            m9 = c(rep(NA, 3)),
                            m10 = c(rep(NA, 3)),
                            m11 = c(rep(NA, 3)),
                            m12 = c(rep(NA, 3)),
                            updated_district_code.y = c(rep(59, 3)),
                            pop = c(rep(107700, 3)),
                            DHF_incidence = c(rep(NA, 3)),
                            log_DHF_incidence = c(rep(NA, 3)),
                            geometry = c(rep(geom_1, 3)),)


# make month 12 data missing
base_model_2001_2006 <- base_model_2001_2006 %>%
  mutate(m12 = case_when(year == 2006 ~ NA))

# rotate month variables into one column
base_model_2001_2006 <- base_model_2001_2006 %>%
  pivot_longer(cols = starts_with("m"), 
               values_to = "m_DHF_cases", 
               names_to = "month")

# take out islands Kien Hai/District + Phu Quoc/City
base_model_2001_2006_1 <- base_model_2001_2006 %>%
  filter(VARNAME_2 != "Kien Hai", 
         VARNAME_2 != "Phu Quoc")


#seasonality_matrix <- model.matrix(~ as.factor(month), data = base_model_2001_2006)

# x_model for seasonality
x_model <- base_model_2001_2006 %>%
  dplyr::select(m_DHF_cases, month)

x_model_1 <- base_model_2001_2006_1 %>%
  dplyr::select(m_DHF_cases, month)

# checking normality of population dist for offset model
hist(MDR_Dengue_map$pop, breaks = 20) 
hist(base_model_2001_2006_1$pop, breaks = 20) 

# set offset model
offset_model <- log(base_model_2001_2006_1$pop)

# arrange by month and then province, then district

# need 72 unique obs for each district --> can put in as missing for ones with less 

# if still not working, can throw out the weirds ones at first
```

## Check Number of rows for each district before running new model
want there to be the same number for every district (12*years)
```{r}

## figuring out problems
for(j in 1: length(unique(base_model_2001_2006_1$VARNAME_2))){
  print(sum(base_model_2001_2006_1$VARNAME_2 == unique(base_model_2001_2006_1$VARNAME_2)[j]))
  if( sum(base_model_2001_2006_1$VARNAME_2 == unique(base_model_2001_2006_1$VARNAME_2)[j]) != 72) {
    print(unique(base_model_2001_2006_1$VARNAME_2)[j])
  }
}



```

Chau Thanh in Hậu Giang Province begins in 2004
Phu Tan in Cà Mau begins in 2004
Dong Hai begins in 2002
Hoa Binh begins in 2006
Binh Thuy begins in 2004
Cai Rang begins in 2004
Ninh Kieu begins in 2004
Thoi Lai is missing 2004-2008
Vinh Thanh begins in 2004
Nam Can begins in 2004
Cao Lanh has a district and city --> 36 years
Chau Thanh A begins in 2004
Long My District begins in 2004
Long My Town beings in 2015
Phung Hiep begins in 2004
Vi Thanh begins in 2004
Vi Thuy begins in 2004
Cu Lao Dung begins in 2003
Nga Nam begins in 2004

## Base Model (based on code from Josh)
```{r}
###############################################################

#Model Fitting

###############################################################

# default of burnin = 10000, n.sample = 110000, thin = 10

results<-ST.CARar(m_DHF_cases ~ 1, data = base_model_2001_2006_1,

                  family = "poisson",

                  W = neighb.mat_2,

                  burnin = 10, 

                  n.sample = 100,

                  thin = 5,

                  prior.mean.beta = rep(0,

                                        times = 1),

                  prior.var.beta = rep((100^2),

                                       times = 1),


                  prior.tau2 = c(0.01, 0.01),
                  
                  AR = 1,

                  MALA = TRUE,

                  verbose = TRUE)

results$summary.results

# want to look at the trace plots for seasonality, overall fitted value (probably won't converge with lower numbers of runs)


```


## Tried Running

```{r}
###############################################################

#Model Fitting

###############################################################
library(CARBayesST)
length(base_model_2001_2006$m_DHF_cases)
length(base_model_2001_2006$month)
length(x_model$m_DHF_cases)
length(neighb.mat)

results<-ST.CARar(m_DHF_cases ~ as.factor(month), 
                  data = base_model_2001_2006_1,

                  family = "poisson",

                  W = neighb.mat_2,

                  burnin = 10, 

                  n.sample = 100,

                  thin = 5,

                  prior.mean.beta = rep(0,

                                        times = ncol(base_model_2001_2006_1) - 1),

                  prior.var.beta = rep((100^2),

                                       times = ncol(base_model_2001_2006_1) - 1),


                  prior.tau2 = c(0.01, 0.01),
                  
                  AR = 1,

                  MALA = TRUE,

                  verbose = TRUE)

results$summary.results

# want to look at the trace plots for seasonality, overall fitted value (probably won't converge with lower numbers of runs)

```
Warning in rnorm(n = length(beta.mean), mean = beta.mean, sd = beta.sd) :
  NAs produced
Warning in rnorm(n = N.all, mean = 0, sd = res.sd) : NAs produced
Warning in matrix(offset, nrow = K, ncol = N, byrow = FALSE) :
  data length [54744] is not a sub-multiple or multiple of the number of rows [132]
Warning in matrix(X.standardised %*% beta, nrow = K, ncol = N, byrow = FALSE) :
  data length [54744] is not a sub-multiple or multiple of the number of rows [132]
Warning in matrix(phi, nrow = K, ncol = N, byrow = FALSE) :
  data length [54744] is not a sub-multiple or multiple of the number of rows [132]
Generating 18 post burnin and thinned (if requested) samples.
  |                                                                                 |   0%Warning in rpois(n = n.miss, lambda = fitted[which.miss == 0]) :
  NAs produced
Warning in matrix(Y.DA, nrow = K, ncol = N, byrow = FALSE) :
  data length [54744] is not a sub-multiple or multiple of the number of rows [132]
Warning in matrix(X.standardised %*% beta, nrow = K, ncol = N, byrow = FALSE) :
  data length [54744] is not a sub-multiple or multiple of the number of rows [132]
Warning in rgamma(1, tau2.shape, scale = (1/tau2.scale)) : NAs produced
Error in if (prob > runif(1)) { : missing value where TRUE/FALSE needed