---
title: "Model Building"
author: "Elisabeth Nelson"
date: "2022-10-20"
output: pdf_document
---
## Load Packages
```{r}
#library(rjags)
library(MCMCpack)
library(stats)
library(CARBayes) # <- this is the package for the model
library(rgdal) # <- for working with shapefiles
library(RColorBrewer) 
library(ggplot2)
library(rgeos) # <- also for working with shapefiles
library(maptools) # shapefiles again
library(spdep) # still shapefiles 
library(ggmap) # plotting shapefiles 
library(sf)
library(dplyr)
library(tidyverse)
library(CARBayesST)

```


## Set-up for Carbays
```{r}
# Number of spatial units (e.g., districts in MDR)
n.district <- 134

# Create an n xô°€ n identity matrix
Id.mat <- diag(n.district)

MDR_2 <- st_read(dsn = '/Users/zoe/Documents/Yale EMD/Rotation 1/Dengue - Vietnam/dengue_exploration/Data/MDR Updated map/VMD_districts_2022.shp')

```

## Neighbor Matrix
```{r}
# Neighbors can either be Queen (any zip that touches another zip - even at 
# corners) or Rook neighbors (share full sides -- not corners)

# The snap dist is governing the NA in neighbor matrix
# Can remove SNAP statement; if Queen = F then single point touch not allowed 
# and Rook distance used.

# When data is missing -- there's more non-zero links involved 
# Try to make snap distance as small as possible while making sure all states 
# have at least one neighbor
neighb <- poly2nb(MDR_2, queen = T, snap = sqrt(0.001))

# Check average number of links -- increase snap distance if it says there is 
# a zip with zero neighbors
neighb # Avg Number of links = 5.19403 
  # regions 74 nand 76 have no links --> islands Kien Hai/District + Phu Quoc/City

# Make neighbor matrix 
# if zero.policy FALSE stop with error for any empty neighbour sets, 
# if TRUE permit the weights list to be formed with zero-length weights vectors
# B is the basic binary coding, W is row standardised (both are commonly used)
neighb.mat <- nb2mat(neighb, zero.policy = T, style = 'B')

    # the island will not have neighbors, so need to exclude it at first
neighb.mat_1 <- neighb.mat[-74, -74]
neighb.mat_2 <- neighb.mat_1[-75, -75]



# For Leroux model, neighbor matrix needs to have 0 if not neighboring, -1 if 
# neighbor, and # of neighbors for each state on diagonal), so...
w.mat <- ifelse(neighb.mat==1,-1,neighb.mat)
num.neigh <- colSums(neighb.mat) # Count the number of neighbors for each state
for (i in 1:nrow(w.mat)) { # Put the number of neighbors on diagonal
  for (j in 1:ncol(w.mat)) {
    if (i==j) {
      w.mat[i,j] <- num.neigh[i]
    }
  }
}
# w.mat <- array(NA, dim=c(n.stzipcode,n.zipcode,n.countries))
# w.mat[,,1] <- ifelse(neighb.mat==1,-1,neighb.mat) # Replace 1 with -1
# num.neigh <- colSums(neighb.mat) # Count the number of neighbors for each state
# for (i in 1:nrow(w.mat[,,1])) { # Put the number of neighbors on diagonal
#   for (j in 1:ncol(w.mat[,,1])) {
#     if (i==j) {
#       w.mat[i,j,1] <- num.neigh[i]
#     }
#   }
# }


```

## Set up Data Frame for Model Training
```{r}
MDR_Dengue_map

# create data frame
base_model_2001_2006 <- fortify(MDR_Dengue_map) %>%
  dplyr::select(VARNAME_2, ENGTYPE_2, district, province, total_DHF_cases, year, m1, m2, m3, m4, m5, m6, m7, m8, m9, m10, m11, m12, updated_district_code.y, pop, DHF_incidence, log_DHF_incidence, geometry) 

base_model_2001_2006 <- base_model_2001_2006 %>% 
  filter(between(year, 2001, 2006))

# make month 12 data missing
base_model_2001_2006 <- base_model_2001_2006 %>%
  mutate(m12 = case_when(year == 2006 ~ NA))

# rotate month variables into one column
base_model_2001_2006 <- base_model_2001_2006 %>%
  pivot_longer(cols = starts_with("m"), 
               values_to = "m_DHF_cases", 
               names_to = "month")

# take out islands Kien Hai/District + Phu Quoc/City
base_model_2001_2006_1 <- base_model_2001_2006 %>%
  filter(VARNAME_2 != "Kien Hai", 
         VARNAME_2 != "Phu Quoc")


#seasonality_matrix <- model.matrix(~ as.factor(month), data = base_model_2001_2006)

# x_model for seasonality
x_model <- base_model_2001_2006 %>%
  dplyr::select(m_DHF_cases, month)

x_model_1 <- base_model_2001_2006_1 %>%
  dplyr::select(m_DHF_cases, month)

# checking normality of population dist for offset model
hist(MDR_Dengue_map$pop, breaks = 20) 
hist(base_model_2001_2006_1$pop, breaks = 20) 

# set offset model
offset_model <- log(base_model_2001_2006_1$pop)

```



## Base Model (based on code from Josh)
```{r}
###############################################################

#Model Fitting

###############################################################


results<-ST.CARar(m_DHF_cases ~ as.factor(month) + offset(offset_model) 
                  + x_model, data = base_model_2001_2006_1,

                  family = "poisson",

                  W = neighb.mat,

                  burnin = 10, 

                  n.sample = 100,

                  thin = 5,

                  prior.mean.beta = rep(0,

                                        times = ncol(x_model_1) - 1),

                  prior.var.beta = rep((100^2),

                                       times = ncol(x_model_1) - 1),


                  prior.tau2 = c(0.01, 0.01),
                  
                  AR = 1,

                  MALA = TRUE,

                  verbose = TRUE)

results$summary.results

# want to look at the trace plots for seasonality, overall fitted value (probably won't converge with lower numbers of runs)

```

## Tried Running

```{r}
###############################################################

#Model Fitting

###############################################################
library(CARBayesST)
length(base_model_2001_2006$m_DHF_cases)
length(base_model_2001_2006$month)
length(x_model$m_DHF_cases)
length(neighb.mat)

results<-ST.CARar(m_DHF_cases ~ as.factor(month), 
                  data = base_model_2001_2006_1,

                  family = "poisson",

                  W = neighb.mat_2,

                  burnin = 10, 

                  n.sample = 100,

                  thin = 5,

                  prior.mean.beta = rep(0,

                                        times = ncol(base_model_2001_2006_1) - 1),

                  prior.var.beta = rep((100^2),

                                       times = ncol(base_model_2001_2006_1) - 1),


                  prior.tau2 = c(0.01, 0.01),
                  
                  AR = 1,

                  MALA = TRUE,

                  verbose = TRUE)

results$summary.results

# want to look at the trace plots for seasonality, overall fitted value (probably won't converge with lower numbers of runs)

```
Warning in rnorm(n = length(beta.mean), mean = beta.mean, sd = beta.sd) :
  NAs produced
Warning in rnorm(n = N.all, mean = 0, sd = res.sd) : NAs produced
Warning in matrix(offset, nrow = K, ncol = N, byrow = FALSE) :
  data length [54744] is not a sub-multiple or multiple of the number of rows [132]
Warning in matrix(X.standardised %*% beta, nrow = K, ncol = N, byrow = FALSE) :
  data length [54744] is not a sub-multiple or multiple of the number of rows [132]
Warning in matrix(phi, nrow = K, ncol = N, byrow = FALSE) :
  data length [54744] is not a sub-multiple or multiple of the number of rows [132]
Generating 18 post burnin and thinned (if requested) samples.
  |                                                                                 |   0%Warning in rpois(n = n.miss, lambda = fitted[which.miss == 0]) :
  NAs produced
Warning in matrix(Y.DA, nrow = K, ncol = N, byrow = FALSE) :
  data length [54744] is not a sub-multiple or multiple of the number of rows [132]
Warning in matrix(X.standardised %*% beta, nrow = K, ncol = N, byrow = FALSE) :
  data length [54744] is not a sub-multiple or multiple of the number of rows [132]
Warning in rgamma(1, tau2.shape, scale = (1/tau2.scale)) : NAs produced
Error in if (prob > runif(1)) { : missing value where TRUE/FALSE needed